What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-07 android-stacks.tex
test
Split lines in 80 characters blocks

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1}, Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2}, Lucas Mariano\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{\{roberta, lucasm\}@dimap.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{gousiosg,a.vandeursen\}@tudelft.nl}}
}

@@

\begin{abstract}
In this work, we perform the first large scale analysis of Java stack traces and pinpoint how they can reveal bad programming practices. We mined and analyzed the stack traces embedded in all issues of Java projects available GitHub. Overall, our research set includes 356,057 issues define at 16,836 projects (from which 28,800 stack traces were extracted). From this set, the stack traces of 482 Android projects were investigated in more detail, in combination with source code and bytecode analysis. In this study some patterns of exception (mis)-use were consistently detected such as: unexpected wrappings (e.g., Errors being wrapped in checked exceptions) - revealing that Java hybrid exception model is not fully used according to its purpose;  undocumented runtime exceptions signaled by third party code - which makes almost impossible for library clients to protect against such exceptions; and a high prevalence of java.lang exceptions reported on issues - representing approximately 50\% of the analyzed issues.

@@

Modern applications have to cope with an increasing number of abnormal computation states that arise as a consequence of faults in the application itself (e.g., access of null references), noisy user inputs or faults in underlying middleware or hardware. The exception handling mechanism~\cite{goodenough1975exception} is one of the most used schemes for detecting and recovering from such exceptional conditions. Although the exception handling mechanism have been embedded in several programming languages (e.g. Java, C++, C\#) and the target of several studies (e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}), the exception handling code is often generally poorly understood and the least tested part of software systems.

@@

%What the stack traces available in these repositories can tell us about the use of exceptions in Java programs?

%To guide the our study we also compile general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate: what stack traces can tell us about the adherence to such practices in Java programs?

To answer these questions, we conduct an exploratory study of stack traces in the wild. Using a custom tool called ExceptionMiner, that was specifically developed to for this study, we mined stack traces from issues across all Java projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects, from which 28,800 stack traces were extracted and used as a source of information for our analysis. The stack trace analysis  was augmented by additional information extracted using byte code and source code analyses done on
a carefully selected subset of our initial sample (482 Android projects).
%This study was divided in two phases: in the fist phase all the issues of all Java projects available on GitHub were analyzed and the embedded stack traces extracted using ExceptionMiner.

To guide this exploration we compiled general guidelines on how to use exceptions proposed by Gosling~\cite{gosling2000java}, Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then we investigate what stack traces can tell us about the adherence to such practices in Java programs.
.%~\todo{This should be our motivation, as this is our biggest contribution. I already changed the text. Lets discuss about this new version}

%GGG: I think this is too much detail for the introduction
%In the second phase of this study, we combined the information of the mined stack traces with source code and bytecode analysis. To do so we selected the subset of 482 Android repositories. The Android repositories were chosen because this kind of application is known for dealing with several sources of exceptions, as they are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and other resources), and multi-htreaded. Moreover, they are based on a single platform which allowed more fine-grained analysis of their exception behaviour.

@@

\begin{itemize}
\item  A multitude undocumented runtime exceptions signaled by third party code (i.e., libraries and Android platform).
\item  Odd exception chains (e.g., Error wrapping checked and runtime exceptions and vice-versa). As an indication that the purpose of Java hybrid exception model may not have been adequately used.
\item  Stack traces caused by uncaught checked exceptions.
\end{itemize}

Hence, the contributions of this study are as follows:
\begin{itemize}
\item  It performs the first large scale analysis of Java stack traces and how they can reveal bad-practices on the use of exceptions in Java. 
\item  It briefly introduces ExceptionMiner, a tool developed to support the analysis.
\end{itemize}

@@

\begin{itemize}
\item  explicitly thrown when a throw statement is executed;
\item  implicitly thrown by the JVM when the evaluation of an expression violates the normal semantics of language (e.g., out-of-bounds array index, division-by-zero, access to a null reference); or 
\item implicitly thrown by the JVM due to an internal error or resource limitation (e.g., OutofMemoryError).
\end{itemize}
%\todo{This sub-section is quite similar to the section introduction, shouldn't we merge them?}
When an exception is signaled it causes a transfer of control from the point where the exception occurred to a point that can be specified by the programmer (exception handler). In Java it is represented by the try-catch block and is responsible for performing any recovery action necessary to bring the software system back to a normal state and, whenever this is not possible, to log the exception and abort the system in an expected safe way.

@@

\subsubsection{Exception Chaining}
Exception Chaining is a common way of propagating exceptions across components of different abstraction levels. It is implemented by special exception constructor which receives another exception as a parameter. An example a multiple exception chaining is depicted on the stack trace in Figure~\ref{fig:wrapping}. 
%This stack trace was extracted from an issue of an Android application in Github.

@@

\subsection{Exception Handling Best Practices}
\label{sec:best}
%\todo{This section feels a bit disconnected.}
Several general guidelines have been proposed on how to use exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward, bloch2008effective}. To guide our investigation, we compile general guidelines related to exception types and the propagation of exception in Java programs \footnote{We also compiled guidelines related to exception handling but they are out of the scope of this paper, since they do not relate to information available on stack traces.}.

@@

\begin{description}
    \item[RQ1] What is the prevalence of stack traces on  reported issues?
    %\item[RQ2] What are the most common exceptions?
    \item[RQ2] What are the common characteristics of such stack traces?
    \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace analysis?
    %~\todo{Perhaps rephrase this to: Are the identified best practices followed in practice?}
\end{description}

@@

%In the second phase of this study we selected a subset of Java applications to be analyzed in more
%detail.The subset chosen was comprised by the Android apps included on GitHub (until 23 Feb 2013).
To identify Android projects, we performed a case insensitive search for the 
term \textsf{android} in the repository's names and short descriptions.
The heuristic filtered 2.542 repositories, from which 589 apps had at least one issue containing a stack trace.

@@

\begin{table}
\centering
\begin{tabular}{p{1em}lrr}
\hline
\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
\hline
   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\  
 \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
%   \multicolumn{2}{l}{\bfseries{ }} &  &  \\   
 \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\  
    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
%\multicolumn{2}{l}{\bfseries{ }} &  &  \\  
\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
    & on issues  & 28,800 & 4,208  \\
    & on issues labeled as defect & 6,529 & 1,699 \\
\hline
\end{tabular}
\caption{Quantitative overview of the analyzed dataset}
\label{tab:overview}
\end{table}

@@

%The results in Table~\ref{tab:overview} are cumulative; however, the numbers
%are not the same among all projects. Figure~\ref{fig:stack traces} 
%presents the descriptive statistics of the range of values of the metrics described in  
%Table~\ref{tab:overview} across all projects in our dataset. 
%There we can see that indeed there is significant variation among projects,
%while the outliers lay at the bottom end of the spectrum, which means that for
%some projects, we can find very limited number of stack traces in their 
%issue reports. Both Java and Android projects feature similar variation,
%which again supports our Android focused analysis later on.

@@

% \begin{itemize}

%     \item What are the most ``popular'' exceptions and what they can tell us?

%     \item Which elements are triggering most of the exceptions and what they 
%     can tell us about the exception structure?

%     \item What is the role of libraries and the underlying platform on crashes?

%     \item What is the role of the application itself on crashes?
% \end{itemize}

% \subsection{The rules of using checked and unchecked exceptions are being obeyed in mobile environment?}

@@

To answer this question, we mined the following information of every extracted stack trace: the top level exception being signaled, its signaler (the first method signature frame after the exception), the root exception (the exception which initiated the stack trace, defined on the bottom of the stack), and the root signaler. The intermediate elements that compose a stack trace as shown in Figure 1 (i.e., the intermediate causes and its execution stacks) were not be considered in the analysis.
%~\todo{Move this to the previous section? IMO, It is better to keep all facts about how we did the research in one place}

@@

%commented only for submission:
%\textbf{We could observe in this study that the NullPointerException and other other implicitly signaled exceptions represented most of the exceptions reported on issues in both Java and Android subset. For such exceptions - which represent programming bugs or resource limitations - there is usually no proper handling besides presenting an error message to the user and restarting the application - only high fault tolerant systems need to provide solutions to handle them [REF: Java Specification].}

@@

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
    \hline
      Programming logic (java.lang and util) & 12625	(41,10\%)	& 2235 (55,64\%)\\
      Resources (IO) & 12440 (40,50\%) & 727 (18,10\%) \\ 
      Security & 220	(0,72\%) & 165 (4,11\%)\\
      Concurrency	& 633 (2,06\%) & 116	(2,89\%)\\
      Backward compatibility & 1580 (5,14\%) & 219 (5,45\%) \\
      Reflection &1413 (4,60\%)   & 91 (2,27\%)\\
      Specific (GUI,FRAMEWORK) &	340	(1,11\%) & 197 (4,90\%)\\
      General (Error, Exception, Runtime) & 1468 (4,78\%) & 267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

@@

%Considering only the packages of each exception we could observe that 81,28\% of all analyzed set of stack traces reported exceptions defined by java.lang (see Table XXX).

%\begin{table}
% \centering
%\begin{tabular}{lrrrr}
%    \hline
% \bfseries{rank} & \bfseries{package} & \bfseries{occurr} &  \bfseries{\%occur} & % \bfseries{projects} \\
%     \hline
%1 &java.lang &	19471 &	81,28\%	& 3080 \\
%2 &	java.io	 & 1222	& 5,10\% & 509 \\
%3 &	java.net &	740	& 3,09\% &	311 \\
%4 &	java.util &	689	& 2,88\% &	287 \\
%5 &	java.lang.reflect & 115 &	0,48\%	& 78 \\
%\hline
%  \end{tabular}
%\caption{Top 5 most popular packages associated to the root exceptions of stack traces}
%\label{tab:toptenpopular2}
%\end{table}

@@

%Even an instance of Throwable as signaled in one project. It points to a the "dont worry from know behaviour".

@@

%In this study we also classified each stack trace according to the type of the root exception been signaled. Table~\ref{typeroottab} presents the types of root causes of all stack traces reported on the Android repositories.

@@

%todo{discomment after submission}
%As we can see from Table ~ref{typeroottab}, that most of the exceptions signaled by third-party code (e.g., Android and libs) were runtime exceptions. When an exception is signaled by third party code, it can be due to 2 reasons: the precondition associated to third party's code was not obeyed; or there is a coding error on third party's code. Only manual code inspection can reveal the reason. However, regardless the reason it is a good practice that all the exceptions signaled by a third part code are properly documented. Manually inspecting a sample of stack traces caused by runtime exceptions thrown by library code (approximately 10\%) we could observe that none of the runtime exceptions signaled by third part code were documented in Javadoc - this analysis did not consider the exceptions implicitly signaled by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a serious problem as it makes very difficult or even impossible for the  client method to protect against  undocumented runtime exceptions signaled by library code. As a consequence, the undocumented runtime exception may remain uncaught leading to  system crashes.

%This result is aligned with the results of other study conducted by from Cabral and Marcques ~\cite{} which also observe that runtime exceptions are most often not documented.

%\textbf{When libraries are used, the developer does not have access to their source code and thus needs to rely on the library documentation about the runtime exceptions that should be thrown – which, more often than not, are neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a consequence, unhandled unchecked exceptions can be seen as one of the major sources of bugs in current Java systems (Jo et al., 2004). }

%On the other hand, we could observe that most of the exceptions signaled by the set of libraries reused by Android platform (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in this work as libcore, were checked exceptions. It is pointed as a good practice for libraries (see Section ~\ref{best}) because by using checked exceptions libraries can define a precise exception interface to its clients. Such libraries are heavily used in several projects, and such precise exception interface may be related to the libraries maturity.

@@

Runtime	&	1101	&	56	&	1446	&	423	&	10	&	3036	\\
Error	&	148	&	34	&	168	&	114	&	9	&	473	\\
Checked	&	115	&	241	&	139	&	176	&	12	&	683	\\
Undefined	&	1	&	0	&	6	&	8	&	0	&	15	\\
Throwable	&	0	&	0	&	1	&	0	&	0	&	1	\\

    \hline
 All  &	1365	&	331	&	1760	&	721	&	31	&	4208	\\
    \hline
  \end{tabular}
\caption{Types of Root Exceptions.}
  \label{fig:typeroottab}
\end{table}

@@

%\subsection{Access to Crash Logs}

%The data

%\subsection{Data Extraction from Natural language Artifacts}

%-Data cleaning removing invalid and duplicated issues.
%-Not every crash information is specified on an issues....

%\subsection{Construct, Internal and External Threats}

@@

\section{Conclusion}
% The line below was meant to be a joke :-)
%Georgious wrote: That's a great paper with profound insights.

@@

%Nowadays many software vendors embed automatic crash reporting tools in their software systems. Hence, whenever a software crashes this tool sends a detailed crash report to its vendors. Moreover, we can also find third party sofware solutions specialized in bug reporting for different kinds of systems specially for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}. There is planty of information to be mined...

>>>>

\author{\IEEEauthorblockN{Roberta Coelho\IEEEauthorrefmark{1},
Georgios Gousios\IEEEauthorrefmark{2}, Arie van Deursen\IEEEauthorrefmark{2},
Lucas Mariano\IEEEauthorrefmark{1}}
\IEEEauthorblockA{\IEEEauthorrefmark{1}Federal University of Rio Grande do Norte\\
Natal, Brazil\\
Email: \texttt{\{roberta, lucasm\}@dimap.ufrn.br}}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Delft University of Technology\\
Delft, The Netherlands\\
Email: \texttt{\{g.gousios,arie.vandeursen\}@tudelft.nl}}
}

@@

\begin{abstract}

@@

Modern applications have to cope with an increasing number of abnormal
computation states that arise as a consequence of faults in the application
itself (e.g., access of null references), noisy user inputs or faults in
underlying middleware or hardware. The exception handling
mechanism~\cite{goodenough1975exception} is one of the most used schemes for
detecting and recovering from such exceptional conditions. Although the
exception handling mechanism have been embedded in several programming languages
(e.g. Java, C++, C\#) and the target of several studies
(e.g.~\cite{miller1997issues,robillard2000designing,shah2010understanding,
garcia2007extracting,garcia2001comparative,cabral2007exception,coelho2011unveiling}),
the exception handling code is often generally poorly understood and the least
tested part of software systems.

@@

%What the stack traces available in these repositories can tell us about the use
%of exceptions in Java programs?

%To guide the our study we also compile general guidelines on how to use
%exceptions proposed by Gosling~\cite{gosling2000java},
%Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
%we investigate: what stack traces can tell us about the adherence to such
%practices in Java programs?

To answer these questions, we conduct an exploratory study of stack traces in
the wild. Using a custom tool called ExceptionMiner, that was specifically
developed to for this study, we mined stack traces from issues across all Java
projects on GitHub. Overall, we analyzed 356,057 issues from 16,836 projects,
from which 28,800 stack traces were extracted and used as a source of
information for our analysis. The stack trace analysis  was augmented by
additional information extracted using byte code and source code analyses done
on a carefully selected subset of our initial sample (482 Android projects).
%This study was divided in two phases: in the fist phase all the issues of all
%Java projects available on GitHub were analyzed and the embedded stack traces
%extracted using ExceptionMiner.

To guide this exploration we compiled general guidelines on how to use
exceptions proposed by Gosling~\cite{gosling2000java},
Wirfs-Brock~\cite{wirfs2006toward} and Bloch~\cite{bloch2008effective} and then
we investigate what stack traces can tell us about the adherence to such
practices in Java programs.
%~\todo{This should be our motivation, as this is
%our biggest contribution. I already changed the text. Lets discuss about this
%new version}

%GGG: I think this is too much detail for the introduction In the second phase
%of this study, we combined the information of the mined stack traces with
%source code and bytecode analysis. To do so we selected the subset of 482
%Android repositories. The Android repositories were chosen because this kind of
%application is known for dealing with several sources of exceptions, as they
%are usually io-intensive (i.e., deals with Network, Bluetooth, camera, and
%other resources), and multi-htreaded. Moreover, they are based on a single
%platform which allowed more fine-grained analysis of their exception behaviour.

@@

\begin{itemize}

  \item  A multitude undocumented runtime exceptions signaled by third party
    code (i.e., libraries and Android platform).

  \item  Odd exception chains (e.g., Error wrapping checked and runtime
    exceptions and vice-versa). As an indication that the purpose of Java hybrid
    exception model may not have been adequately used.

  \item  Stack traces caused by uncaught checked exceptions.

\end{itemize}

Hence, the contributions of this study are as follows:
\begin{itemize}

  \item  It performs the first large scale analysis of Java stack traces and how
    they can reveal bad-practices on the use of exceptions in Java.

  \item  It briefly introduces ExceptionMiner, a tool developed to support the
    analysis.

\end{itemize}

@@

\begin{itemize} \item  explicitly thrown when a throw statement is executed;
  \item  implicitly thrown by the JVM when the evaluation of an expression
    violates the normal semantics of language (e.g., out-of-bounds array index,
    division-by-zero, access to a null reference); or \item implicitly thrown by
    the JVM due to an internal error or resource limitation (e.g.,
    OutofMemoryError).  \end{itemize}
%\todo{This sub-section is quite similar to the section introduction, shouldn't
%we merge them?}
When an exception is signaled it causes a transfer of control from the point
where the exception occurred to a point that can be specified by the programmer
(exception handler). In Java it is represented by the try-catch block and is
responsible for performing any recovery action necessary to bring the software
system back to a normal state and, whenever this is not possible, to log the
exception and abort the system in an expected safe way.

@@

\subsubsection{Exception Chaining} Exception Chaining is a common way of
propagating exceptions across components of different abstraction levels. It is
implemented by special exception constructor which receives another exception as
a parameter. An example a multiple exception chaining is depicted on the stack
trace in Figure~\ref{fig:wrapping}.
%This stack trace was extracted from an issue of an Android application in
%Github.

@@

\subsection{Exception Handling Best Practices}
\label{sec:best}

%\todo{This section feels a bit disconnected.}
Several general guidelines have been proposed on how to use
exceptions~\cite{mandrioli1992advances,gosling2000java,wirfs2006toward,
bloch2008effective}. To guide our investigation, we compile general guidelines
related to exception types and the propagation of exception in Java programs
\footnote{We also compiled guidelines related to exception handling but they are
out of the scope of this paper, since they do not relate to information
available on stack traces.}.

@@

\begin{description}

  \item[RQ1] What is the prevalence of stack traces on  reported issues?
    %\item[RQ2] What are the most common exceptions?

  \item[RQ2] What are the common characteristics of such stack traces?

  \item[RQ3] What patterns of exception (mis)-use emerge from the stack trace
    analysis?
    %~\todo{Perhaps rephrase this to: Are the identified best practices followed
    %in practice?}

\end{description}

@@

%In the second phase of this study we selected a subset of Java applications to
%be analyzed in more detail.The subset chosen was comprised by the Android apps
%included on GitHub (until 23 Feb 2013).
To identify Android projects, we performed a case insensitive search for the
term \textsf{android} in the repository's names and short descriptions.  The
heuristic filtered 2.542 repositories, from which 589 apps had at least one
issue containing a stack trace.

@@

\begin{table}
\centering
\begin{tabular}{p{1em}lrr}
\hline
\multicolumn{2}{l}{\bfseries{Issue-level metrics}} & \bfseries{Java} & \bfseries{Android Subset}\\
\hline
   \multicolumn{2}{l}{\bfseries{Repositories}} & 16,836 & 482\\
 \multicolumn{2}{l}{\bfseries{Issues}} & 356,057 & 32,582\\
    & incl. exceptions & 30,236 (8,5\%) & 3,956 (12,1\%) \\
    & incl. stack traces & 21,013 (5,9\%) & 3,101 (7,3\%) \\
 \multicolumn{2}{l}{\bfseries{Defect Issues}} & 55,226 & 2,376 \\
    & incl. exception & 6,741 (12,2\%) & 1,557 (65,6\%) \\
    & incl. stack trace &  5,196 (9,4\%) & 1,393 (58,7\%)\\
\multicolumn{2}{l}{\bfseries{Exception Stack Traces}} &  & \\
    & on issues  & 28,800 & 4,208  \\
    & on issues labeled as defect & 6,529 & 1,699 \\
\hline
\end{tabular}
\caption{Quantitative overview of the analyzed dataset}
\label{tab:overview}
\end{table}

@@

@@

@@

To answer this question, we mined the following information of every extracted
stack trace: the top level exception being signaled, its signaler (the first
method signature frame after the exception), the root exception (the exception
which initiated the stack trace, defined on the bottom of the stack), and the
root signaler. The intermediate elements that compose a stack trace as shown in
Figure 1 (i.e., the intermediate causes and its execution stacks) were not be
considered in the analysis.
%~\todo{Move this to the previous section? IMO, It is better to keep all facts
%about how we did the research in one place}

@@

%commented only for submission:
\textbf{We could observe in this study that the NullPointerException and other
other implicitly signaled exceptions represented most of the exceptions reported
on issues in both Java and Android subset. For such exceptions - which represent
programming bugs or resource limitations - there is usually no proper handling
besides presenting an error message to the user and restarting the application -
only high fault tolerant systems need to provide solutions to handle them [REF:
Java Specification].}

@@

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \hline
    \bfseries{Category} & \bfseries{Java} & \bfseries{Android Subset} \\
    \hline
      Programming logic (java.lang and util) & 12625  (41,10\%) & 2235 (55,64\%)\\
      Resources (IO)                         & 12440 (40,50\%)  & 727 (18,10\%) \\
      Security                               & 220  (0,72\%)    & 165 (4,11\%)\\
      Concurrency                            & 633 (2,06\%)     & 116  (2,89\%)\\
      Backward compatibility                 & 1580 (5,14\%)    & 219 (5,45\%) \\
      Reflection                             & 1413 (4,60\%)    & 91 (2,27\%)\\
      Specific (GUI,FRAMEWORK)               & 340 (1,11\%)     & 197 (4,90\%)\\
      General (Error, Exception, Runtime)    & 1468 (4,78\%)    & 267 (6,65\%)\\
    \hline
  \end{tabular}
  \caption{Characterization of the 100 root causes.}
  \label{tab:tophundrend}
\end{table}

@@

%Considering only the packages of each exception we could observe that 81,28\%
%of all analyzed set of stack traces reported exceptions defined by java.lang
%(see Table XXX).

%\begin{table}
% \centering
%\begin{tabular}{lrrrr}
%    \hline
% \bfseries{rank} & \bfseries{package} & \bfseries{occurr} &  \bfseries{\%occur} & % \bfseries{projects} \\
%     \hline
%1 & java.lang         & 19471 & 81,28\% & 3080 \\
%2 & java.io           & 1222  & 5,10\%  & 509 \\
%3 & java.net          & 740   & 3,09\%  & 311 \\
%4 & java.util         & 689   & 2,88\%  & 287 \\
%5 & java.lang.reflect & 115   & 0,48\%  & 78 \\
%\hline
%  \end{tabular}
%\caption{Top 5 most popular packages associated to the root exceptions of stack traces}
%\label{tab:toptenpopular2}
%\end{table}

@@

%Even an instance of Throwable as signaled in one project. It points to a the
%"dont worry from know behaviour".

@@

%In this study we also classified each stack trace according to the type of the
%root exception been signaled. Table~\ref{typeroottab} presents the types of
%root causes of all stack traces reported on the Android repositories.

@@

%todo{discomment after submission} As we can see from Table ~ref{typeroottab},
%that most of the exceptions signaled by third-party code (e.g., Android and
%libs) were runtime exceptions. When an exception is signaled by third party
%code, it can be due to 2 reasons: the precondition associated to third party's
%code was not obeyed; or there is a coding error on third party's code. Only
%manual code inspection can reveal the reason. However, regardless the reason it
%is a good practice that all the exceptions signaled by a third part code are
%properly documented. Manually inspecting a sample of stack traces caused by
%runtime exceptions thrown by library code (approximately 10\%) we could observe
%that none of the runtime exceptions signaled by third part code were documented
%in Javadoc - this analysis did not consider the exceptions implicitly signaled
%by JVM (e.g. NullPoiterException). The undocumented runtime exceptions are a
%serious problem as it makes very difficult or even impossible for the  client
%method to protect against  undocumented runtime exceptions signaled by library
%code. As a consequence, the undocumented runtime exception may remain uncaught
%leading to  system crashes.

%This result is aligned with the results of other study conducted by from Cabral
%and Marcques ~\cite{} which also observe that runtime exceptions are most often
%not documented.

%\textbf{When libraries are used, the developer does not have access to their
%source code and thus needs to rely on the library documentation about the
%runtime exceptions that should be thrown – which, more often than not, are
%neither complete nor precise (Thomas, 2002; Sacramento et al., 2006). As a
%consequence, unhandled unchecked exceptions can be seen as one of the major
%sources of bugs in current Java systems (Jo et al., 2004). }

%On the other hand, we could observe that most of the exceptions signaled by the
%set of libraries reused by Android platform (i.e., org.apache.harmony,
%org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml, javax.) referred in
%this work as libcore, were checked exceptions. It is pointed as a good practice
%for libraries (see Section ~\ref{best}) because by using checked exceptions
%libraries can define a precise exception interface to its clients. Such
%libraries are heavily used in several projects, and such precise exception
%interface may be related to the libraries maturity.

@@

Runtime   & 1101 & 56  & 1446 & 423 & 10 & 3036  \\
Error     & 148  & 34  & 168  & 114 & 9  & 473 \\
Checked   & 115  & 241 & 139  & 176 & 12 & 683 \\
Undefined & 1    & 0   & 6    & 8   & 0  & 15  \\
Throwable & 0    & 0   & 1    & 0   & 0  & 1 \\
    \hline
All       &	1365	&	331	&	1760	&	721	&	31	&	4208	\\
    \hline
  \end{tabular}
\caption{Types of Root Exceptions.}
  \label{fig:typeroottab}
\end{table}

@@

@@

\section{Conclusion}

@@

%Nowadays many software vendors embed automatic crash reporting tools in their
%software systems. Hence, whenever a software crashes this tool sends a detailed
%crash report to its vendors. Moreover, we can also find third party sofware
%solutions specialized in bug reporting for different kinds of systems specially
%for the increasing marked of mobile apps~\cite{BugSe14,BugSn14,Googl14,Acra14}.
%There is planty of information to be mined...
